Verifying a Binary Search Tree with Fine-Grained Locking

### 2020-09-25 concurrency/Verification of Concurrent Binary Search Tree.tex

Linearizability is one of the most popular strong correctness condition for the most of concurrent data-structure, which implies that every operation appears to take place atomically, in some order, consistent with the real-time ordering of those operations.

\subsection{Locking and atomicity} 
In order to record the action each thread performed as a linear history, we will define global invariants using the ghost state, which are similar to lock invariants but are not associated with any particular memory location. Instead, a global invariant is true before and after every step of a program, acting as a publicly accessible resource. A program instruction can use the contents of global invariant if it can guarantee that no one will ever see an intermediate state in which the invariant does not hold. 
\subsection{Insert and Lookup}

>>>>

Linearizability is one of the most popular strong correctness condition for the most of concurrent data-structure, which implies that every operation appears to take place atomically, in some order, consistent with the real-time ordering of those operations. 
In order to record the action each thread performed as a linear history, we will define global invariants using the ghost state, which are similar to lock invariants but are not associated with any particular memory location. Instead, a global invariant is true before and after every step of a program, acting as a publicly accessible resource. A program instruction can use the contents of global invariant if it can guarantee that no one will ever see an intermediate state in which the invariant does not hold.

\subsection{Fine-grained locking and atomicity}

To prove that an operation on a data structure satisfies a logically atomic specification, we must show that there are no visible intermediate states of the operation, i.e., that other threads see the data structure as unchanged until the linearization point at which the operation takes effect. This is often implemented with either a lock-free series of atomic memory accesses, in which case all but the last access must make changes that are considered ``invisible'', or a coarse-grained lock, in which case any changes may be made and no other thread can access the data structure until the operation is complete. Fine-grained locking presents an interesting middle ground between these two approaches: other threads may continue to access the data structure as long as they do not require a section that is currently locked, and the visible changes may be implemented with multiple non-atomic operations in the locked section.

In an Iris-style separation logic, the lock-free approach means that each atomic operation gets access to the state of the data structure stored in an atomic shift, and must either maintain its previous state or (at the linearization point) move to a new one; in the coarse-grained approach we can store the state of the data structure in a lock invariant, and the function has access to all of it throughout the critical section. With fine-grained locking, neither of these approaches quite applies as-is: there are no atomic operations by which we access the state of the data structure, but neither is there a lock whose invariant controls access to the entire structure. However, we can take a modified form of the lock-free approach, where we treat each critical section as an atomic operation that can access the atomic shift. Typically, we will perform a sequence of non-atomic operations in the body of the critical section without access to the global state of the data structure, and then (before exiting the critical section) access the atomic shift and show that we have either maintained the original state or satisfied the postcondition of the operation. In this section, we describe a general approach to associating fine-grained locks with pieces of the abstract state of a data structure, allowing them to interact with atomic shifts and be used to prove atomic specifications. (Note that this approach is not novel to us; it is adapted from something Ralf Jung worked on. We need to be more clear about the relationship.)

technical material: definition of sync\_inv, sync\_commit and sync\_rollback

\subsection{Insert and Lookup}
